\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[hidelinks]{hyperref}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{0pt}

\title{Solving Fredholm Integral Equations by Writing Code:\\
\large A Program-of-Thought Experiment on \texttt{test\_100\_v2}}
\author{Fred-LLM project}
\date{13 July 2026}

\begin{document}
\maketitle

\begin{abstract}
\noindent
Frontier models are better programmers than they are algebraists, so we stop
asking them to do algebra. Each equation in the benchmark is re-rendered as a
short block of executable SymPy definitions, and the model is asked for a
Python script that solves it. We run that script in a sandbox with a 60-second
timeout and one repair round, and score whatever it prints with the
repository's unchanged evaluation pipeline. On the 100-equation
\texttt{test\_100\_v2} set this reaches 42/100 with gpt-5.4, the best result
we have from any GPT-family method: the plain-prompt baseline scores 35 and
the five-agent agentic workflow scores 38 at 2.3 times the cost. On
gpt-5.4-mini the method scores 35/100, which matches the plain gpt-5.4
baseline at about a quarter of its price. An ablation that shows the model the
same code-shaped prompt but takes its answer directly, without execution,
gains only 3 points over the baseline; execution is what pays, not the
representation. The method has one systematic failure: it finds almost none of
the 15 no-solution items. Most scripts do attempt a singularity check, but
they choose tolerances that a float system at these magnitudes never trips. The fine-tuned 360M specialist from the
previous report keeps the overall lead (46 plain, 53 with solver routing),
and it wins exactly where the scripts lose, on the decision items.
\end{abstract}

\section{Motivation}
The July~7 agentic pilot bought its accuracy with orchestration: five method
agents per equation, a SymPy verifier, and a repair round, at \$10.62 per
hundred equations on gpt-5.4. The July~12 fine-tuning report then showed that
a deterministic degenerate-kernel solver, called at the right moments, solves
more of this benchmark than any GPT prompt. Both results point the same way.
The hard part of a separable-kernel equation is not choosing the method; it is
carrying out pages of exact arithmetic on coefficients like
$-797089.486$. Our working hypothesis was that models can set this
arithmetic up but cannot carry it out.

Program-of-thought prompting (PAL, PoT) splits the two: the model writes the
setup as code, an interpreter does the arithmetic. This experiment tests that
split on Fredholm equations, and uses an ablation to check where any gain
comes from.

\section{Method}

\subsection{Code representation}
\texttt{experiments/code\_prompt/build\_prompts.py} converts each
\texttt{test\_100\_v2} item (kernel, right-hand side, $\lambda$, domain, all
taken from the same metadata the basic prompts use) into executable
definitions:

\begin{verbatim}
from sympy import *
x, t = symbols('x t')
lam = 0.6800510922329444
a, b = -1.1508025669664157, 9.746297944109074
K = sin(x)*cos(t)      # kernel K(x, t)
f = ...                # right-hand side f(x)
# u(x) - lam * Integral(K * u(t), (t, a, b)) = f
\end{verbatim}

LaTeX-to-SymPy conversion goes through the repository's
\texttt{parse\_latex\_to\_sympy}; all 100 items convert, and every generated
block was executed locally as a sanity check before any API call. Items with
$\lambda = 0$ are rendered as first-kind equations, the same rule the basic
renderer applies, so the prompts leak no ground truth.

\subsection{Two variants}
The \texttt{code} variant shows the model this block and asks for an answer
directly, in the standard \texttt{SOLUTION:} format. It isolates the effect of
seeing the equation as code rather than as LaTeX.

The \texttt{code\_exec} variant asks for one self-contained Python script that
solves the equation and prints the same three-line format, with the solution
in LaTeX via \texttt{sympy.latex()}. A new wrapper,
\texttt{CodeExecModelRunner} (\texttt{src/llm/code\_exec\_runner.py}, enabled
by a \texttt{model.code\_exec} config section), extracts the script, runs it
in an isolated subprocess (\texttt{python -I}, 60-second timeout), and feeds
one failure back to the model for a single repair round. If the script still
fails, the raw model text is scored instead. The script's stdout replaces the
model's response, so the existing parser, evaluator, and metrics apply without
modification. This is the same wrapping pattern the agentic runner uses, which
keeps every number in this report comparable to the earlier ones.

Both prompts state the degenerate-kernel recipe (reduce to a linear system
for the integral coefficients; singular and consistent means a family,
singular and inconsistent means no solution). The agentic workflow's
method directives say the same thing, so no method receives hints the others
lack.

\subsection{Runs}
Temperature 0.1, \texttt{max\_tokens} 4096, and the evaluation tolerances of
the July~7 pilot (series 0.01, approximate-coefficient 0.001, regularized
0.001), all through the same router. An 8-item pilot ran first, one item per
solution type plus one extra, at \$0.11 for both variants together. The full
runs cost \$0.76 (mini, \texttt{code}), \$0.78 (mini, \texttt{code\_exec}),
and \$4.71 (gpt-5.4, \texttt{code\_exec}). We skipped the gpt-5.4
\texttt{code} run: the mini ablation had already answered the representation
question, and the extra \$3 would not have changed any conclusion.

\section{Results}

\begin{table}[h]
\centering
\begin{tabular}{llrrr}
\toprule
Model & Method & Correct/100 & Type acc. & Cost \\
\midrule
gpt-5.4-mini & basic prompt            & 21 & 0.253 & --- \\
gpt-5.4-mini & agentic multi-method    & 32 & 0.310 & --- \\
gpt-5.4-mini & \texttt{code} (no execution) & 24 & 0.256 & \$0.76 \\
gpt-5.4-mini & \texttt{code\_exec}     & \textbf{35} & 0.198 & \$0.78 \\
\midrule
gpt-5.4 & basic prompt                 & 35 & 0.385 & \$3.01 \\
gpt-5.4 & agentic multi-method         & 38 & 0.420 & \$10.62 \\
gpt-5.4 & \texttt{code\_exec}          & \textbf{42} & 0.244 & \$4.71 \\
\midrule
SmolLM2-360M & plain SFT               & 46 & 0.680 & local \\
SmolLM2-360M & plain SFT + solver routing & \textbf{53} & 0.680 & local \\
(no LM) & separable solver always      & 28 & 0.300 & free \\
\bottomrule
\end{tabular}
\caption{\texttt{test\_100\_v2} results. Baseline, agentic, and SmolLM2 rows
are from the July~7 and July~12 reports. ``Correct/100'' counts unparseable
responses as wrong.}
\end{table}

\texttt{code\_exec} beats the basic prompt by 14 points on mini and 7 on
gpt-5.4, and beats the agentic workflow on both models while using one model
call per equation instead of five to ten. The representation-only ablation
gains 3 points. Reading the equation as code barely matters; running the
model's code is the mechanism.

\begin{table}[h]
\centering
\begin{tabular}{lrrrrrrr}
\toprule
Method & exact & approx & series & discrete & family & regul. & none \\
\midrule
mini, basic     & 6/30  & 3/14 & 0/10 & 0/9 & 10/10 & 0/10 & 2/15 \\
mini, agentic   & 13/30 & 2/15 & 0/10 & 0/10 & 10/10 & 2/10 & 5/15 \\
mini, \texttt{code}      & 9/30  & 0/15 & 0/9  & 0/9 & 10/10 & 0/10 & 5/15 \\
mini, \texttt{code\_exec} & \textbf{20/28} & 3/13 & 0/10 & 0/7 & 10/10 & 2/10 & 0/15 \\
gpt-5.4, \texttt{code\_exec} & \textbf{23/29} & 2/14 & 0/9 & 0/8 & 10/10 & \textbf{6/10} & 1/15 \\
\bottomrule
\end{tabular}
\caption{Per-type correct counts. Denominators below the type's size mean
some responses for that type failed to parse.}
\end{table}

The exact-symbolic column carries the result. The basic mini prompt solves 6
of 30; the same model writing scripts solves 20 of 28. gpt-5.4 reaches 23 of
29, and 6 of its 10 first-kind items, the best regularized score of any GPT
method. The scripts also keep the family items at 10/10 because the models
special-case $f = 0$ in code and report the resonant family.

Script reliability was high on both models. On mini, 92 of 100 scripts ran
on the first try and one response was prose with no code block; the repair
round fixed 5 of the remaining 7, and the two that stayed broken were a
60-second timeout and a persistent traceback. On gpt-5.4, 87 ran first try
and the repair round fixed 9 of 13, leaving 4 items scored on raw text.
Repairs are the difference between the 100 prompts and the 107 and 113
requests the cost tracker recorded.

\section{Why execution wins, and where it fails}
Solving $u(x) = f(x) + \lambda \sum_i c_i g_i(x)$ requires integrating
products of the $h_i$ against $f$ and inverting a small matrix with
coefficients that run to six significant figures times $10^5$. The baseline
responses at temperature 0.1 return only a final answer, so we cannot see
where the internal derivation goes wrong; what the results show is that the
same model that gets 6 (mini) or 13 (agentic mini) of the exact items right
on its own gets 20 right once an interpreter does the arithmetic. The
generated scripts are not trivial, the median is 70 lines, but every line is
setup: define the kernel factors, build the coefficient matrix, hand it to
SymPy. The ability being measured shifts from symbolic manipulation to
problem setup, and setup saturates earlier. Mini plus an interpreter equals
gpt-5.4 alone, at 26\% of its cost.

The failure is just as mechanical, and it is not for lack of trying. Every
no-solution item in this benchmark is a resonance case: $\lambda$ sits at an
eigenvalue, the linear system is singular, and the inhomogeneous part is
inconsistent with it. 73 of the 100 gpt-5.4 scripts, and 50 of the mini
ones, do attempt a rank, determinant, or conditioning check. The checks
fail anyway: one representative gpt-5.4 script computes at 80-digit
precision and then tests \texttt{rref} pivots against a $10^{-28}$
threshold, which no near-singular system with coefficients of order $10^5$
will ever cross. The float system looks regular, \texttt{linsolve} returns a
huge answer, and the script prints it with \texttt{HAS\_SOLUTION: yes}. Mini
lost all 15 such items, gpt-5.4 lost 14. Choosing a numerically meaningful
tolerance turns out to be the one part of the setup the models cannot do
either. Type accuracy drops for a related reason: a script that computed a
float answer tends to label it \texttt{exact\_symbolic} regardless of what
the item actually is.

This is why the fine-tuned specialist still leads. The evaluator scores 35 of
the 100 items purely on decisions (no-solution, family, regularized), and
decisions are what the SFT model learned and what scripts are worst at. The
two failure profiles barely overlap, which suggests a cheap hybrid: let
something else own the no-solution decision (the SFT router, or a
conditioning check whose tolerance the harness fixes instead of the model),
and let the script own the algebra. The solver-routing
result from the previous report (46 to 53 from routing alone) is the same
shape of fix.

\section{Comparison with the agentic workflow}
Agentic buys its lift with parallel method agents, a residual verifier, and
2.2 times the requests of a baseline run. \texttt{code\_exec} spends one call
plus a subprocess and gains more on both models. The two are orthogonal
rather than competing: the agentic verifier could score script outputs, and a
script is in effect a fifth method agent whose work is checked by an
interpreter instead of a judge. The scaffolding-helps-weaker-models pattern
from the agentic pilot repeats here, 14 points of lift on mini against 7 on
gpt-5.4, at roughly triple the magnitude the agentic workflow managed.

\section{Limitations}
Single run per configuration at temperature 0.1, so the numbers carry
sampling noise we have not measured; the 3-point representation effect in
particular is within that noise. The first-kind items are named as first-kind
in a code comment, which is one step more explicit than the basic prompt's
rendering (the form is visible in both, but the baseline model has to
recognize it); part of the 6/10 regularized score may be that labeling.
Series and discrete-point items score zero under every GPT method, including
this one, and scripts do not change that: those items need a different output
contract, not better arithmetic. The sandbox is a subprocess with a timeout,
not a hardened jail; that is fine for SymPy scripts from a model solving math
and would not be fine for arbitrary code. And the killed first attempt at the
gpt-5.4 run cost an unrecoverable partial batch, since the
\texttt{code\_exec} path has no checkpointing; long runs should add the
per-equation checkpoint the agentic runner already has.

\section{Reproducing}
\begin{verbatim}
.venv/bin/python experiments/code_prompt/build_prompts.py
.venv/bin/python -m src.cli run --config configs/test_100v2_gpt54_code_exec.yaml
.venv/bin/python experiments/code_prompt/compare.py
\end{verbatim}
Artifacts, including every generated script and its execution record, are in
\texttt{results/code\_prompt\_2026-07-13/}. The narrative version of this
report is \texttt{experiments/code\_prompt/REPORT.md}.

\end{document}
