\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[hidelinks]{hyperref}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{0pt}

\title{Four Ways to Solve a Fredholm Equation with an LLM:\\
\large Baseline Prompting vs.\ Agentic vs.\ Code Execution vs.\ Fine-Tuning, Side by Side}
\author{Fred-LLM project}
\date{23 July 2026}

\begin{document}
\maketitle

All numbers are on the same benchmark: \texttt{test\_100\_v2}, 100 Fredholm
equations of the second kind across seven solution types (30 exact symbolic,
15 approximate-coefficient, 15 no-solution, 10 series, 10 discrete-point, 10
solution-family, 10 regularized first-kind), scored by the same evaluation
pipeline. Counts treat unscored responses as wrong. Full details are in the
July 7 (agentic), July 12 (fine-tuning), and July 13 (code execution) reports.

\section{The four methods}

\textbf{Baseline.} One zero-shot prompt with the equation in LaTeX; the model
answers directly in the structured \texttt{SOLUTION} format. One model call
per equation, temperature 0.1.

\textbf{Agentic multi-method.} Five specialist agents (degenerate kernel,
Adomian, Neumann series, Fredholm alternative, numerical) attempt each
equation in parallel. A deterministic SymPy check substitutes each candidate
back into the equation; residual $<10^{-6}$ counts as verified, one repair
round re-prompts failures, and a rule-based selector picks the winner
(smallest residual, else majority vote, else best remaining). Five to seven
model calls per equation.

\textbf{Code execution (program of thought).} The equation is re-rendered as
executable SymPy definitions and the model writes one Python script that
solves it and prints the standard output format. The script runs in a
sandboxed subprocess (60-second timeout, one repair round) and its stdout is
scored as if the model had said it. One model call per equation. An ablation
that shows the code-shaped prompt without executing anything gains only 3
points, so execution, not representation, is the mechanism.

\textbf{Fine-tuning + solver routing.} SmolLM2-360M-Instruct, fully
fine-tuned on 11{,}100 equations from the project's augmentation pipeline,
run locally. Plain text labels beat a special-token variant of the same
training (0.505 vs.\ 0.370 evaluator accuracy), so plain is shown. The
routing variant adds no training: when the model declares an equation
solvable, a deterministic degenerate-kernel solver computes the answer and
replaces the model's; the model's type decision only routes. The solver
without the router scores 28/100, because it happily ``solves'' equations
that have no solution.

\section{Results by solution type}

\begin{table}[h]
\centering
\small
\setlength{\tabcolsep}{4.5pt}
\begin{tabular}{llccccccccr}
\toprule
 & & exact & approx & series & discr. & family & regul. & none & \textbf{total} & cost \\
Model & Method & /30 & /15 & /10 & /10 & /10 & /10 & /15 & \textbf{/100} & (\$) \\
\midrule
gpt-5.4-mini & baseline    & 6  & 3  & 0 & 0 & 10 & 0  & 2 & 21 & 0.57 \\
gpt-5.4-mini & agentic     & 13 & 2  & 0 & 0 & 10 & 2  & 5 & 32 & 5.12 \\
gpt-5.4-mini & code exec.  & 20 & 3  & 0 & 0 & 10 & 2  & 0 & 35 & 0.78 \\
\midrule
gpt-5.4 & baseline         & 21 & 1  & 0 & 0 & 10 & 1  & 2 & 35 & 3.01 \\
gpt-5.4 & agentic          & 24 & 0  & 0 & 0 & 10 & 2  & 2 & 38 & 10.62$^{*}$ \\
gpt-5.4 & code exec.       & 23 & 2  & 0 & 0 & 10 & \textbf{6} & 1 & 42 & 4.71 \\
\midrule
SmolLM2-360M & plain SFT   & 7  & \textbf{11} & 0 & 0 & 10 & \textbf{10} & \textbf{8} & 46 & local \\
SmolLM2-360M & SFT + routing & 18 & 7 & 0 & 0 & 10 & \textbf{10} & \textbf{8} & \textbf{53} & local \\
\bottomrule
\end{tabular}
\caption{Correct counts per solution type on \texttt{test\_100\_v2}. Cost is
API spend per 100-equation run (list-price proxy). $^{*}$Partial log after a
checkpoint resume; the full run is estimated near \$36.}
\end{table}

Read by column, the table splits into three regimes.

\textbf{Algebra (exact symbolic, 30 items).} Scaffolding is everything for
weak models and marginal for strong ones: the mini goes $6 \to 13 \to 20$
across baseline, agentic, and code execution, while gpt-5.4 moves only
$21 \to 24 \to 23$. The 360M model cannot do the algebra itself (7/30) but
recovers to 18/30 by routing to the deterministic solver.

\textbf{Decisions (none, regularized, family --- 35 items scored on
classification).} Fine-tuning owns this regime: 10/10 regularized and 8/15
none, against 0--2 for nearly every GPT configuration. The one GPT exception
is gpt-5.4 code execution on regularized items (6/10). Code execution is
worst on no-solution items (0--1 of 15): the scripts attempt singularity
checks but pick tolerances a float system never trips, so they print a huge
``solution'' instead of abstaining. Family items are free for everyone
(10/10 everywhere).

\textbf{Numeric types (approx-coef, series, discrete --- 35 items).}
Series and discrete-point are zero for every method; they need a different
output contract, not better solving. Approximate-coefficient is the SFT
model's other win (11/15 plain), since it learned the expected output format;
GPT methods manage at most 3.

\section{Cost}

Per 100 equations: the baseline costs \$0.57 (mini) to \$3.01 (gpt-5.4);
code execution costs about the same as a baseline (\$0.78, \$4.71) because it
is still one call per equation; agentic costs $5$--$7\times$ more (\$5.12,
$\approx$\$36 full-run estimate). Per \emph{extra} correct answer over the
same-model baseline: agentic pays \$0.41 on the mini but $\approx$\$11 on
gpt-5.4; code execution pays \$0.02 on the mini and \$0.24 on gpt-5.4. The
fine-tuned model has no per-query API cost at all (a few hours of training on
a laptop), and holds the overall lead at 53/100.

\section{Takeaway}

Code execution is the best way to spend API money on this benchmark (42/100
at baseline-level cost); agentic orchestration only pays where the base model
is weak; and a 360M local specialist plus a deterministic solver beats every
GPT configuration (53/100) by winning the decision items the GPTs throw away.
The failure profiles barely overlap --- scripts own the algebra, the
specialist owns the decisions --- so the obvious next step is the hybrid.
(gpt-5.5 repeats the gpt-5.4 pattern, 36 baseline / 38 agentic, and is
omitted from the table.)

\end{document}
